---
title: "Problem set 1"
author: "YOUR NAME HERE"
date: "DATE HERE"
---

# Task 1

Tell me that you watched the short videos and completed the series of R Markdown tutorials.


# Task 2

## Life expectancy in 2007

```{r load-packages, warning=FALSE, message=FALSE}
# Notice the warning=FALSE and message=FALSE in the chunk options. R spits out 
# a lot of messages when you load tidyverse and we don't want those in the 
# final document.

library(tidyverse)  # This loads ggplot2, dplyr, and other packages you'll need
library(gapminder)  # This loads the gapminder data
```

Let's first look at the first few rows of data:

```{r view-data}
head(gapminder)
```

Right now, the `gapminder` data frame contains rows for all years for all countries. We want to only look at 2007, so we create a new data frame that filters only rows for 2007.

Note how there's a weird sequence of characters: `%>%`. This is called a *pipe* and lets you chain functions together. We could have also written this as `gapminder_2007 <- filter(gapminder, year == 2007)`.

```{r filter-2007}
gapminder_2007 <- gapminder %>%
  filter(year == 2007)

head(gapminder_2007)
```

Now we can plot a histogram of 2007 life expectancies with the default settings:

```{r plot-2007-1}
ggplot(data = gapminder_2007, 
       mapping = aes(x = lifeExp)) +
  geom_histogram()
```

R will use 30 histogram bins by default, but that's not always appropriate, and it will yell at you for doing so. **Adjust the number of bins to 2, then 40, then 100.** **What's a good number for this data? Why?**

TYPE YOUR ANSWER HERE.

```{r plot-2007-2}
ggplot(data = gapminder_2007, 
       mapping = aes(x = lifeExp)) +
  geom_histogram(bins = 2)
```

## Average life expectancy in 2007 by continent

We're also interested in the differences of life expectancy across continents. First, we can group all rows by continent and calculate the mean

This is where the `%>%` function is actually super useful. Remember that it lets you chain functions together—this means we can read these commands as a set of instructions: take the `gapminder` data frame, filter it, group it by continent, and summarize each group by calculating the mean. Without using the `%>%`, we could write this same chain like this: `summarize(group_by(filter(gapminder, year == 2007), continent), avg_life_exp = mean(lifeExp))`. But that's *awful* and impossible to read and full of parentheses that can easily be mismatched.

```{r calc-mean}
gapminder_cont_2007 <- gapminder %>%
  filter(year == 2007) %>% 
  group_by(continent) %>%
  summarize(avg_life_exp = mean(lifeExp))

head(gapminder_cont_2007)
```

Let's plot these averages as a bar chart:

```{r plot-2007-bar}
ggplot(data = gapminder_cont_2007, 
       mapping = aes(x = continent, y = avg_life_exp, fill = continent)) + 
  geom_col()
```

Then, let's plot them as density distributions. We don't need to use the summarized data frame for this, just the original filtered `gapminder_2007` data frame:

```{r plot-2007-density}
ggplot(data = gapminder_2007, 
       mapping = aes(x = lifeExp, fill = continent)) + 
  geom_density()
```

Now let's plot life expectancies as violin charts. These are the density distributions turned sideways:

```{r plot-2007-violin}
ggplot(data = gapminder_2007, 
       mapping = aes(x = continent, y = lifeExp, fill = continent)) + 
  geom_violin()
```

Finally, we can add actual points of data for each country to the violin chart:

```{r plot-2007-violin-points}
ggplot(data = gapminder_2007, 
       mapping = aes(x = continent, y = lifeExp, fill = continent)) + 
  geom_violin() +
  geom_point()
```

The bar chart, density plot, violin plot, and violin plot + points each show different ways of looking at a single number—the average life expectancy in each continent. **Answer these questions:**

- Which plot is most helpful? TYPE YOUR ANSWER HERE.
- Which ones show variability? TYPE YOUR ANSWER HERE.
- What's going on with Oceania? TYPE YOUR ANSWER HERE.


# Task 3: R and ggplot

```{r load-libraries-1, warning=FALSE, message=FALSE}
# Technically you don't need to run this because we loaded tidyverse earlier in Task 2
library(tidyverse)
```

## 3.2.4

**1: Run `ggplot(data = mpg)`. What do you see?**

```{r blank-plot, fig.width=2, fig.height=2}
ggplot(data = mpg)
```

TYPE YOUR ANSWER HERE. (hint: I gave you the answer in the chunk name)

(Notice how I used `fig.width` and `fig.height` in the chunk options. You can click on the little gear icon in the far left of the chunk to change other options.)

**2: How many rows are in `mpg`? How many columns?**

```{r mpg-details}
nrow(mpg)
ncol(mpg)

# Or
dim(mpg)

# Or
mpg
```

**3: What does the `drv` variable describe? Read the help for `?mpg` to find out.**

TYPE YOUR ANSWER HERE.

**4: Make a scatterplot of `hwy` vs `cyl`.**

```{r hwy-cyl-scatterplot}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))
```

**5: What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?**

```{r class-drv-scatterplot}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = class, y = drv))
```

TYPE YOUR ANSWER HERE.


## 3.3.1

Add your own chunks here and answer the questions.

## 3.5.1

Add your own chunks here and answer the questions.

## 3.6.1

Add your own chunks here and answer the questions.

## 3.8.1

Add your own chunks here and answer the questions.
